### 2025-07-20 cluster-analysis.qmd
20250708 - track chunk execution times
20250717 - chapter overviews
20250720 - conclusion sections

@@

Based on this analysis (and the variables included), there appear to be three types of Wide Receivers.
Type 1 Wide Receivers includes the Elite WR1s who are strong possession receivers (note: not all players in a given cluster map on perfectly to the typology—i.e., not all Type 1 Wide Receivers are elite WR1s).
They tend to have the lowest drop percentage, the shortest average depth of target, and the fewest receiving air yards per reception.
They tend to have the most receptions per season and break the most tackles.

Type 2 Wide Receivers includes the consistent contributor, WR2 types.
They had fewer receptions and fewer broken tackles than Type 1 Wide Receivers.
Their average depth of target was longer than Type 1, and they had more receiving air yards per reception than Type 1.

Type 3 Wide Receivers includes the deep threats.
They have the greatest average depth of target and the most receiving yards per reception.
However, they also have the fewest receptions, the highest drop percentage, and the fewest broken tackles.
Thus, they may be considered the boom-or-bust Wide Receivers.

@@

>>>>

This chapter provides an overview of cluster analysis.

@@

Based on this analysis (and the variables included), there appear to be three types of Wide Receivers.
We examined the following variables: the player's vertical jump in the NFL Combine,40-yard-dash time in the NFL Combine, height, weight, average depth of target, drop percentage, receiving air yards per reception, broken tackles, and receptions per season.

Type 1 Wide Receivers included the Elite WR1s who are strong possession receivers (note: not all players in a given cluster map on perfectly to the typology—i.e., not all Type 1 Wide Receivers are elite WR1s).
They tended to have the lowest drop percentage, the shortest average depth of target, and the fewest receiving air yards per reception.
They tended to have the most receptions per season and break the most tackles.

Type 2 Wide Receivers included the consistent contributor, WR2 types.
They had fewer receptions and fewer broken tackles than Type 1 Wide Receivers.
Their average depth of target was longer than Type 1, and they had more receiving air yards per reception than Type 1.

Type 3 Wide Receivers included the deep threats.
They had the greatest average depth of target and the most receiving yards per reception.
However, they also had the fewest receptions, the highest drop percentage, and the fewest broken tackles.
Thus, they may be considered the boom-or-bust Wide Receivers.

@@

The goal of cluster analysis is to identify distinguishable subgroups of people.
There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), and others.
The present chapter used model-based clustering to identify tiers of players based on projected points.
Using various performance metrics of Wide Receivers, we identified three subtypes of Wide Receivers: 1) Elite WR1s who are strong possession receivers; 2) Consistent Contributor/WR2s; 3) deep threats/boom-or-bust receivers.
The "Elite WR1s" tended to have the lowest drop percentage, the shortest average depth of target, the fewest receiving air yards per reception, the most receptions per season, and the most broken tackles.
The "Consistent Contributor/WR2s" had fewer receptions and fewer broken tackles than the Elite WR1s; their average depth of target was longer than Elite WR1s, and they had more receiving air yards per reception than Elite WR1s.
The "Deep Threat/Boom-or-Bust" receivers had the greatest average depth of target and the most receiving yards per reception; however, they also had the fewest receptions, the highest drop percentage, and the fewest broken tackles.
In sum, cluster analysis can be a useful way of identifying subgroups of individuals who are more similar to one another on various characteristics.

### 2025-07-20 decision-making.qmd
20250720 - conclusion sections

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/projectionsWithActuals_seasonal.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/projectionsWithActuals_weekly.RData", fsep = ""))
#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_seasonal.RData", fsep = ""))
#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
```

```{r}
load(file = "./data/projectionsWithActuals_seasonal.RData")
load(file = "./data/projectionsWithActuals_weekly.RData")
#load(file = "./data/players_projections_seasonal.RData")
#load(file = "./data/players_projections_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
load(file = "./data/player_stats_weekly.RData")
```

@@

Averaging projections from individuals tends to yield predictions that are more accurate than the accuracy of most forecasters in the group [@Mannes2014].
Indeed, when at least some of of the projections bracket the truth, averaged predictions must be more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—and averaged predictions are often much more accurate [@Larrick2024].
When referring to the accuracy of the "average individual forecaster", we are referring to accuracy in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—not to the accuracy of the forecaster at the 50th percentile.
If none of the projections bracket the truth—e.g., all projections overestimate the truth—averaged predictions will be as accurate than the average individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError) [@Larrick2024].
In sum, "Averaging the answers of a crowd, therefore, ensures a level of accuracy no worse than the average member of the crowd and, in some cases, a level better than nearly all members" [@Larrick2024, p. 126].
Moreover, averaged projections tend to be more accurate than consensus-based judgments from groups of people that interact and discuss, due to cognitive biases associated with the social interaction among groups, such as herding in which people align their behavior with others [@Mannes2014; @Simoiu2019], though discussion can be helpful in some contexts [@Larrick2024].

@@

The relatively modest accuracy of the projections by so-called fantasy "experts'" and of their average of their projections could occur for a number of reasons.
One possibility is that the level of expertise of the "expert" forecasters in terms of being able to provide accurate forecasts is not strong.
That is, because football performance and injuries are so challenging to predict, individual forecasters' projections may not be particularly close to the truth.

@@

Proportion of players whose bracketing rate was zero:

@@

@@

@@

Because fantasy football is not 100% luck, it is not the same as a slot machine.
However, there are still unpredictable elements (e.g., injuries), and there is considerable luck involved.
As a result, sports betting and gambling on fantasy football still involve components of [variable ratio reinforcement](#sec-reinforcementIntermittent) that lend them to being potentially highly addictive.

@@

>>>>

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/projectionsWithActuals_seasonal.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/projectionsWithActuals_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
```

```{r}
load(file = "./data/projectionsWithActuals_seasonal.RData")
load(file = "./data/projectionsWithActuals_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
load(file = "./data/player_stats_weekly.RData")
```

@@

Averaging projections from individuals tends to yield predictions that are more accurate than the accuracy of most forecasters in the group [@Mannes2014].
Indeed, when at least some of the projections bracket the truth, averaged predictions must be more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—and averaged predictions are often much more accurate [@Larrick2024].
When referring to the accuracy of the "average individual forecaster", we are referring to accuracy in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—not to the accuracy of the forecaster at the 50th percentile.
If none of the projections bracket the truth—e.g., all projections overestimate the truth—averaged predictions will be as accurate than the average individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError) [@Larrick2024].
In sum, "Averaging the answers of a crowd, therefore, ensures a level of accuracy no worse than the average member of the crowd and, in some cases, a level better than nearly all members" [@Larrick2024, p. 126].
Moreover, averaged projections tend to be more accurate than consensus-based judgments from groups of people that interact and discuss, due to cognitive biases associated with the social interaction among groups, such as herding in which people align their behavior with others [@Mannes2014; @Simoiu2019], though discussion can be helpful in some contexts [@Larrick2024].

@@

The relatively modest accuracy of the projections by so-called fantasy "experts" and of their average of their projections could occur for a number of reasons.
One possibility is that the level of expertise of the "expert" forecasters in terms of being able to provide accurate forecasts is not strong.
That is, because football performance and injuries are so challenging to predict, individual forecasters' projections may not be particularly close to the truth.

@@

Below is the proportion of players whose bracketing rate was zero:

@@

```{r}
#| include: false

proportionZeroBracketingRate_seasonal <- seasonalIndividualSources %>% 
  filter(!is.na(bracketingRate)) %>% 
  mutate(isZeroBracketingRate = ifelse(bracketingRate == 0, 1, 0)) %>% 
  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n())
```

A high proportion (`{r} petersenlab::apa(proportionZeroBracketingRate_seasonal * 100, decimals = 0)`%) of players' forecasts did not bracket the truth.

@@

```{r}
#| include: false

proportionZeroBracketingRate_weekly <- weeklyIndividualSources %>% 
  filter(!is.na(bracketingRate)) %>% 
  mutate(isZeroBracketingRate = ifelse(bracketingRate == 0, 1, 0)) %>% 
  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n())
```

A high proportion (`{r} petersenlab::apa(proportionZeroBracketingRate_weekly * 100, decimals = 0)`%) of players' forecasts did not bracket the truth.

@@

Because fantasy football is not 100% luck, it is not the same as a slot machine.
However, there are still unpredictable elements (e.g., injuries), and there is considerable luck involved.
As a result, sports betting and gambling on fantasy football still involve components of [variable ratio reinforcement](#sec-reinforcementIntermittent) that lend them to being potentially highly addictive despite people's tendency to lose money.

@@

The wisdom of the crowd is the idea that the average of the predictions of many people is often more accurate than the prediction of individual experts.
When at least some of the projections bracket—i.e., fall on opposite sides of—the truth, averaged predictions are more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE).
Crowd-averaged projections tend to be most accurate when the crowd consists of individuals who hold expertise in the domain such that they will make predictions that fall close to the truth, there is relatively low variability in the expertise of the individual forecasters in terms of their ability to make accurate forecasts, there is cognitive diversity among the forecasters, the projections are made independently—i.e., the forecasters are not aware of others’ forecasts and do not discuss or interact with the other forecasters, the bracketing rate is high, and there are at least 5–10 sources of projections.
In the case of fantasy football, however, a high proportion of players' forecasts do not bracket the truth, suggesting that the so-called experts are not accurate in predicting players' future fantasy performance.
Nevertheless, going with the crowd-averaged projections tends to be more accurate than relying on an individual projection source.
Performance in fantasy football is around half and half luck versus skill.
Sports betting and gambling on fantasy football involve components of [variable ratio reinforcement](#sec-reinforcementIntermittent) that lend them to being potentially highly addictive.

### 2025-07-20 factor-analysis.qmd
20250717 - chapter overviews
20250720 - conclusion sections

@@

>>>>

This chapter provides an overview of factor analysis, which encompasses a range of latent variable modeling approaches, including [exploratory](#sec-efa) and [confirmatory](#sec-cfa) factor analysis.

@@

Factor analysis is a class of latent variable models that aims to identify the optimal, parsimonious latent structure for a group of variables.
Latent variables are ways of studying and operationalizing theoretical constructs that cannot be directly observed or quantified.
Factor analysis encompasses two general types: [confirmatory factor analysis](#sec-cfa) and [exploratory factor analysis](#sec-efa).
[*Exploratory factor analysis*](#sec-efa) (EFA) is used when the researcher has no a priori hypotheses about how a set of variables is structured.
[*Confirmatory factor analysis*](#sec-cfa) (CFA) is used when a researcher wants to evaluate how well a hypothesized model fits.
A goal of factor analysis is to balance accuracy (i.e., variance accounted for) and parsimony (i.e., simplicity).
Factor analysis estimates the latent factors as the common variance among the variables that load onto that factor and discards the remaining variance as "error".
There are many [decisions to make in factor analysis](#sec-factorAnalysisDecisions).
These decisions can have important impacts on the resulting solution.
Thus, it can be helpful for theory and interpretability to help guide decision-making when conducting factor analysis.
Using both [exploratory](#sec-efa) and [confirmatory](#sec-cfa) factor analysis, we were able to identify three latent factors that accounted for considerable variance in the variables we examined, pertaining to Quarterbacks: 1) usage; 2) aggressiveness; 3) performance.
We were then able to determine which players were highest and lowest on each of these factors.

### 2025-07-20 machine-learning.qmd
20250720 - conclusion sections

The most accurate model, in terms of $R^2$, had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
This suggests that the other predictors examined did not have [incremental validity](#sec-incrementalValidity)—they did not show utility above and beyond a player's fantasy points in predicting their future fantasy points.
If anything, considering the additional predictor variables showed *decremental* validity—the inclusion of the additional variables worsened prediction, possibly by weakening the predictive association of the previous season's fantasy points and misattributing its predictive effect to other, related predictors in the model.

@@

The most accurate model in terms of [mean absolute error](#sec-meanAbsoluteError) was the [tree-boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`.
However, even in the most accurate model, the predicted values were over `{r} petersenlab::apa(mae_minimum, decimals = 0)` points away from the actual values, suggesting that the model-predicted values were not particularly accurate.
In sum, the [regression model with one predictor](#sec-regressionOnePredictor) was the most accurate in terms of differentiating between players ($R^2$), whereas the [tree-boosting model](#sec-treeBoosting) was the most accurate in terms of how close the predicted values were to the actual values.

@@

>>>>

The most accurate model, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
This suggests that the other predictors examined did not have [incremental validity](#sec-incrementalValidity)—they did not show utility above and beyond a player's fantasy points in predicting their future fantasy points.
If anything, considering the additional predictor variables showed *decremental* validity—the inclusion of the additional variables worsened prediction, possibly by weakening the predictive association of the previous season's fantasy points and misattributing its predictive effect to other, related predictors in the model.

@@

The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`.
However, even in the most accurate model, the predicted values were over `{r} petersenlab::apa(mae_minimum, decimals = 0)` points away from the actual values, suggesting that the model-predicted values were not particularly accurate.
In sum, the [regression model with one predictor](#sec-regressionOnePredictor) was the most accurate in terms of differentiating between players ($R^2$), whereas the [tree-boosting model](#sec-treeBoosting) was the most accurate in terms of how close the predicted values were to the actual values.

@@

Machine learning attempts to maximize prediction accuracy.
There are various types of machine learning models.
[Supervised learning](#sec-machineLearningTypesSupervised) involves learning from data where the correct classification or outcome is known.
[Unsupervised learning](#sec-machineLearningTypesUnsupervised) involves learning from data without known classifications.
Machine learning can incorporate many predictor variables, so it is important to perform [cross-validation](#sec-crossValidation) to avoid [overfitting](#sec-overfitting).
We examined multiple machine learning models to identify the model that generates the most accurate prediction of Quarterbacks' future fantasy points.
All models explained less than half of the variance in players' fantasy points.
The most accurate model, in terms of [discrimination](#sec-discrimination), was achieved using regression with one predictor: the player's fantasy points in the previous season.
The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which accounted for the longitudinal nature of the data.
The considerable unexplained variance suggests that we are missing key variables that are important predictors of future fantasy performance.

### 2025-07-20 modern-portfolio-theory.qmd
20250719 - citations
20250720 - conclusion sections

The tangency portfolio is the portfolio with the highest Sharpe ratio (i.e., the highest ratio of return to risk).
In other words, it is the portfolio with the greatest risk-adjusted returns.

@@

In summary, fantasy football is similar to stock picking.
You are most likely to pick the best players if you go with the [wisdom of the crowd](#sec-wisdomOfTheCrowd) (e.g., average projections across projection sources) and diversify.
Most projections are public information, so you might wonder whether using crowd projections gains you anything because everybody else has access to public information.
However, this is also the case with stocks, and people still consistently perform best over time when they go with the market.
Nevertheless, crowd projections are not highly accurate.
And fantasy football is a game, so feel free to have fun and deviate from the crowd!
However, you may be just as (if not more) likely to be wrong by deviating from the crowd.

>>>>

The tangency portfolio is the portfolio with the highest Sharpe ratio—i.e., the highest ratio of return to risk.
In other words, it is the portfolio with the greatest risk-adjusted returns.

@@

In summary, fantasy football is similar to stock picking.
You are most likely to pick the best players if you go with the [wisdom of the crowd](#sec-wisdomOfTheCrowd) (e.g., average projections across projection sources) and diversify.
Players—like stocks—have expected returns and risk associated with them.
You can think of constructing your team like constructing a stock portfolio, where your goal is to gain the maximum expected returns for a given level of risk.
The greater risk you are willing to take, the greater the potential returns, but alse the lower the potential downside if the players underperform.
Thus, you may want your team to be composed of a diverse combination of some higher risk players and some lower risk players.
We demonstrate the [efficient frontier](#sec-efficientFrontier) as a valuable tool for identifying the combination of players with the maximum expected returns relative to their risk.
Most projections are public information, so you might wonder whether using crowd projections gains you anything because everybody else has access to public information.
However, this is also the case with stocks, and people still consistently perform best over time when they go with the market.
Nevertheless, crowd projections are not highly accurate.
And fantasy football is a game, so feel free to have fun and deviate from the crowd!
However, you may be just as (if not more) likely to be wrong by deviating from the crowd.

### 2025-07-20 pca.qmd
20250712 - tweaks
20250720 - conclusion sections

@@

We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
The optimal number of components based on VSS criterion is three components.

@@

>>>>

This chapter provides an overview of principal component analysis as a useful technique for data reduction.

@@

We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
The optimal number of components based on the VSS criterion is three components.

@@

Principal component analysis (PCA) is a technique used for data reduction—reducing a large set of a variables down to a smaller set of components that capture most of the variance in the larger set.
There are many [decisions to make in factor analysis](#sec-pcaDecisions).
These decisions can have important impacts on the resulting solution.
Thus, it can be helpful for theory and interpretability to help guide decision-making when conducting factor analysis.
There are several differences between [factor analysis](#sec-factorAnalysis) and PCA.
Unlike [factor analysis](#sec-factorAnalysis), which estimates the latent factors as the common variance among the variables that load onto that factor and discards the remaining variance as "error", PCA uses all variance of variables and assumes variables have no error.
Thus, PCA does not account for measurement error.
Using PCA, we were able to identify three PCA components that accounted for considerable variance in the variables we examined, pertaining to Quarterbacks: 1) usage; 2) aggressiveness; 3) performance.
We were then able to determine which players were highest and lowest on each of these components.

### 2025-07-20 simulation.qmd
20250708 - track chunk execution times
20250717 - chapter overviews
20250720 - conclusion sections

@@

>>>>

This chapter provides an overview of various approaches to simulation, including [bootstrapping](#sec-simulationProjectedPointsBootstrapping) and the [Monte Carlo](#sec-simulationOverviewMonteCarlo) method.

@@

A simulation is an "imitative representation" of a phenomenon that could exist the real world.
In statistics, simulations are computer-driven investigations to better understand a phenomenon by studying its behavior under different conditions.
Statistical simulations can be conducted in various ways.
Two common types of simulations are [bootstrapping](#sec-simulationOverviewBootstrapping) and [Monte Carlo](#sec-simulationOverviewMonteCarlo) simulation.
[Bootstrapping](#sec-simulationOverviewBootstrapping) involves repeated resampling (with replacement) from observed data.
[Monte Carlo](#sec-simulationOverviewMonteCarlo) simulation involves repeated random sampling from a known distribution.
We demonstrated [bootstrapping](#sec-simulationOverviewBootstrapping) and [Monte Carlo](#sec-simulationOverviewMonteCarlo) approaches to simulating the most likely range of outcomes for a player in terms of fantasy points.

### 2025-07-20 sports-cognitive-psychology.qmd
20250717 - chapter overviews
20250720 - conclusion sections

@@

>>>>

This chapter provides an overview of some of the roles of cognitive psychology and analytics in sports.

@@

Football has recently seen greater use of analytics; however, its uptake has been slower compared to other sports like baseball and basketball.
Examples of changes in football due to the use of analytics include [more often going for it on fourth down](#sec-coachingGoingForIt), a greater emphasis on the passing game, drafting Running Backs later in the draft (and, more generally, valuing Running Backs less), and trading down in the draft to obtain more low picks (rather than having fewer high picks) because top picks are frequently overvalued.
Historically, despite the decision hurting their chances of winning, teams rarely went for it on fourth down and elected to punt or kick a field goal instead, likely due to [loss aversion bias](#sec-cognitiveBiasesLossAversion).
[Loss aversion bias](#sec-cognitiveBiasesLossAversion) may lead coaches to make more conservative play-calling, in general, such as running the ball when passing would tend to be more advantageous.
Rates of going for it on fourth down in the NFL were low until 2017.
After the 2017 season, rates of going for it on fourth down increased dramatically.
In addition to poor decisions relating to going for it on fourth down, there is not strong evidence that [icing the kicker](#sec-icingTheKicker) works, suggesting that it may better suit your team to use timeouts in other situations.

There are many [psychological factors that influence sports performance](#sec-psychologicalFactorsPerformance).
According to the Yerkes–Dodson law that describes the association between arousal and performance.
For simple tasks, performance tends to increase with arousal.
There is an optimal range of arousal for complex and challenging tasks, such as those in sports, which often require divided attention and multitasking.
Too little arousal is bad because it leads to insufficient attention on the task at hand; however, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking.
When a task is highly practiced or automatic, excessive attention can hurt performance.
By contrast, for tasks that require conscious control, decision-making, and real-time adjustments, more focused attention tends to lead to better performance.

### 2025-07-20 time-series-analysis.qmd
20250708 - track chunk execution times
20250720 - conclusion sections

@@

>>>>

This chapter provides an overview of time series analysis.

@@

That is, time series analysis seeks to evaluate change over time to predict future values.
There many different types of time series analyses.
We demonstrated use of [autoregressive integrated moving average](#sec-timeSeriesARIMA) (ARIMA) models to predict future fantasy points.
ARIMA models aim to describe how how earlier levels of a variable are correlated with later levels of the same variable.
ARIMA models perform best when there is a clear pattern where later values are influenced by earlier values.
ARIMA models assume that the data are stationary (i.e., there are no long-term trends), are non-seasonal (i.e., there is no consistency of the timing of the peaks or troughs in the line), and that earlier values influence later values.
This may not strongly be the case in fantasy football, so ARIMA models may not be particularly useful in forecasting fantasy football performance.
We also used [Bayesian mixed models](#sec-timeSeriesBayesian) to generate forecasts of future performance and plots of individuals model-implied performance by age and position.
